### Driver hourly revenue

Primary Factor: Sustainability

**Description:** Driver Hourly Revenue reflects whether the passenger transportation sector is producing a sustainable level of revenue for drivers to promote retention of their workforce. It is important to note that hourly revenue does not reflect take-home pay, which depends on factors such as business-level labour payment arrangements, employment classification (e.g., employee vs. independent contractor), operating costs, and hours worked. However, revenue per hour remains a useful indicator of market stability and service sustainability from a labour supply perspective.

**Measurement:** This indicator captures the median[1] of the total fare[2] collected by taxi and TNS drivers per working hour in a defined geographical area for over a month. It reflects whether the passenger transportation sector is producing a sustainable level of revenue for drivers to promote retention of their workforce.

**Interpretation:** Higher hourly revenue indicates stronger revenue generation for the sector, supporting its long-term operational sustainability. This indicator measures median hourly revenue at the sector level and does not reflect individual driver take-home earnings, which can vary based on a variety of factors. The focus is on the sector’s overall financial health and its ability to maintain service levels over time.


Policy Context: Public Need and Accessibility Factor

\newpage

#### `r region_name` Regional District taxi and TNS Driver hourly revenue Findings 


`r tbl_caption(paste(region_name, "Regional District taxi and TNS Driver hourly revenue Findings"))`


```{r}
make_indicator_table_compact(
  data      = raw_norm,
  indicator = "DRIVER_EARNING",
  region    = region_name,
  date_from = date_from,
  date_to   = date_to,
  digits    = 2,
  currency  = TRUE)
```

`r fig_caption(paste(region_name, "Regional District taxi and TNS Driver hourly revenue Findings"))`

```{r, fig.width=10, fig.height=5}
chart <- make_indicator_chart_lines(
data       = raw_norm,
indicator  = "DRIVER_EARNING",  # Adjust indicator if needed
region     = region_name,
services   = c("TAXI", "TNS"),
area_type  = "REGIONAL",
date_from  = date_from,
date_to    = date_to,
title      = "Taxi and TNS Driver hourly revenue($)",
colors     = c(TAXI = "#69B7FF", TNS = "#1E3A8A")
)
chart

```

In `r end_label`, `r region_name` Regional District recorded a total of `r metric("DRIVER_EARNING","TAXI","last_value_fmt")` in driver earnings for taxi services. On a year-over-year (YoY) basis, average driver earnings for taxi services were `r metric("DRIVER_EARNING","TAXI","yoy_fmt")`while on a year-to-date (YTD) basis it was `r metric("DRIVER_EARNING","TAXI","ytd_yoy_fmt")` relative to the same period in `r prev_year`. On a month-over-month (MoM) basis, driver earnings for taxi trips changed by `r metric("DRIVER_EARNING","TAXI","mom_fmt")`. `r trend_narrative("DRIVER_EARNING", "TAXI")` `r shift_narrative("DRIVER_EARNING", "TAXI")`

In `r end_label`, the `r region_name` Regional District recorded a total of `r metric("DRIVER_EARNING","TNS","last_value_fmt")` in driver earnings for TNS services. Compared with `r prev_label`, average driver earnings for TNS services were `r metric("DRIVER_EARNING","TNS","yoy_fmt")` (YoY). Year-to-date through `r end_label`, earnings for TNS services were `r metric("DRIVER_EARNING","TNS","ytd_yoy_fmt")` relative to the same period in `r prev_year`. On a month-over-month (MoM) basis, driver earnings for TNS trips changed by `r metric("DRIVER_EARNING","TNS","mom_fmt")`. `r trend_narrative("DRIVER_EARNING", "TNS")` `r shift_narrative("DRIVER_EARNING", "TNS")`

**Limitations:**

## Variety Factor (SEC)
There is no analysis related to this factor.