### Driver hourly revenue

Primary Factor: Sustainability

**Description:** Driver Hourly Revenue reflects whether the passenger transportation sector is producing a sustainable level of revenue for drivers to promote retention of their workforce. It is important to note that hourly revenue does not reflect take-home pay, which depends on factors such as business-level labour payment arrangements, employment classification (e.g., employee vs. independent contractor), operating costs, and hours worked. However, revenue per hour remains a useful indicator of market stability and service sustainability from a labour supply perspective.

**Measurement:** This indicator captures the median[1] of the total fare[2] collected by taxi and TNS drivers per working hour in a defined geographical area for over a month. It reflects whether the passenger transportation sector is producing a sustainable level of revenue for drivers to promote retention of their workforce.

**Interpretation:** Higher hourly revenue indicates stronger revenue generation for the sector, supporting its long-term operational sustainability. This indicator measures median hourly revenue at the sector level and does not reflect individual driver take-home earnings, which can vary based on a variety of factors. The focus is on the sector’s overall financial health and its ability to maintain service levels over time.


Policy Context: Public Need and Accessibility Factor

?????

#### `r region_name` Regional District Taxi and TNS Driver hourly revenue Findings 


*`r tbl_caption(paste(region_name, "Regional District Taxi and TNS Driver hourly revenue Findings"))`*


```{r}
tbl <- make_indicator_table_compact(
  data       = raw_norm,
  indicator  = "DRIVER_EARNING",
  region     = region_name,
  services   = c("TAXI","TNS"),
  area_type  = "REGIONAL",
  date_from  = date_from,
  date_to    = date_to,
  digits     = 2
)
tbl
```

*`r fig_caption(paste(region_name, "Regional District Taxi and TNS Driver hourly revenue Findings"))`*

```{r, fig.width=10, fig.height=5}
chart <- make_indicator_chart_lines(
data       = raw_norm,
indicator  = "DRIVER_EARNING",  # Adjust indicator if needed
region     = region_name,
services   = c("TAXI", "TNS"),
area_type  = "REGIONAL",
date_from  = date_from,
date_to    = date_to,
title      = "Taxi and TNS Driver hourly revenue($)",
colors     = c(TAXI = "#69B7FF", TNS = "#1E3A8A")
)
chart

```
