## Accessibility Factor (PN)

### Number of Wheelchair Accessible Vehicles 

**Primary Factor:** Accessibility

**Description:**The number of wheelchair accessible vehicles can be further divided into two sub-metrics: (1) the number of active wheelchair accessible vehicles and (2) the number of active wheelchair accessible vehicles providing at least one accessible trip. The number of active wheelchair accessible vehicles (WAVs) refers to the total number of vehicles equipped to accommodate passengers who use wheelchairs and that have been in operation in an area over a month. On the other hand, the number of active wheelchair accessible vehicles (WAVs) providing at least one accessible trip refers to the number of active WAVs that have recorded at least one accessible trip in the Trip Database.

**Measurement:** The number of active WAVs is determined by cross-referencing the list of WAVs reported to the Registrar’s Office during licence renewal with the list of active vehicles (which may not necessarily have provided accessible trips) reported to the Trip Database. Then, the number of active wheelchair accessible vehicles providing at least one accessible trip is determined by limiting the list of active WAVs to those reported as having provided accessible trips. Both metrics can be further normalized per 1,000 persons to indicate their growth relative to the population.

**Interpretations:** The number of active WAVs indicates the accessibility and inclusiveness of passenger transportation services for individuals who use wheelchairs. Generally, a higher number of active WAVs, especially compared to growth in overall demand (proxied by the number of WAVs per 1,000 persons), indicates a greater level of availability for these individuals. A lack of growth in the number of WAVs (both sub-metrics) would indicate suppliers’ unwillingness to match the demand, potentially due to the high costs associated with WAVs. 

The Board’s public need policy in the Policy Manual lists examples of the accessibility  factor, which are set out below alongside their corresponding implications on the number of active wheelchair vehicles:

- **The proposed service provides more or improved service to persons with mobility disabilities, including WAVs.**  
  -There are people who require access to, or are seeking access to, or would likely access a new, expanded, or improved passenger transportation service.
  - The number of WAVs on the road directly and positively correlates with the extent and quality of service provided to individuals with mobility disabilities.
  
#### `r region_name` Regional District Number of Wheelchair Accessible Vehicles Findings



`r tbl_caption(paste(region_name, "Regional District Number of Active Wheelchair Accessible Vehicles"))`
```{r}
make_indicator_table_compact(
  data      = raw_norm,
  indicator = "ACTIVE_WAV",
  region    = region_name,
  date_from = date_from,
  date_to   = date_to,
  digits    = 0,
  currency  = FALSE)
```

`r tbl_caption(paste(region_name, "Regional District Number of Active Wheelchair Accessible Vehicles Providing at Least One Accessible Trip "))`

```{r}
make_indicator_table_compact(
  data      = raw_norm,
  indicator = "ACTIVE_WAV_WITH_ACCESSIBLE_TRIP",
  region    = region_name,
  date_from = date_from,
  date_to   = date_to,
  digits    = 0,
  currency  = FALSE)

```

<!-- `r tbl_caption(paste(region_name, "Regional District Number of 1Active Wheelchair Accessible Vehicles per 1,000 Persons"))` -->


<!-- ```{r} -->
<!-- make_new_table( -->
<!--   data      = raw_norm, -->
<!--   indicator = "ACTIVE_WAV_PER_1000_PERSONS", -->
<!--   region    = region_name, -->
<!--   date_from = date_from, -->
<!--   date_to   = date_to, -->
<!--   font_size     = 8, -->
<!--   page_width_in = 8,    -->
<!--   year_col_in   = 0.4 -->
<!-- ) -->

<!-- ``` -->

<!-- Note: The above table reports the number of WAVs per 1,000 persons and the annual population data is available from BC Stats [^2] -->

<!-- [^2]:BC Stats. (n.d.). Population Estimates & Projections for British Columbia. Retrieved January 20, 2025, from https://bcstats.shinyapps.io/popApp/ -->
<!-- `r tbl_caption(paste(region_name, "Regional District Number of Active Wheelchair Accessible Vehicles Providing at Least One Accessible Trip per 1,000 Persons"))` -->


<!-- ```{r} -->
<!-- make_new_table( -->
<!--   data      = raw_norm, -->
<!--   indicator = "ACTIVE_WAV_WITH_ACCESS_TRIP_PER_1000_PERSONS", -->
<!--   region    = region_name, -->
<!--   date_from = date_from, -->
<!--   date_to   = date_to, -->
<!--   font_size     = 8, -->
<!--   page_width_in = 8,    -->
<!--   year_col_in   = 0.4 -->
<!-- ) -->
<!-- ``` -->
<!-- Note: The above table reports the number of WAVs providing accessible trips per 1,000 persons and the annual population data is available from BC Stats[^3] -->

<!-- [^3]:BC Stats. (n.d.). Population Estimates & Projections for British Columbia. Retrieved January 20, 2025, from https://bcstats.shinyapps.io/popApp/ -->

`r fig_caption(paste(region_name, "Regional District Trip Volume"))`
```{r, fig.width=10, fig.height=5}
chart <- make_multi_indicator_chart_lines(
  data       = raw_norm,
  indicators = c(
    "ACTIVE_WAV",
    "ACTIVE_WAV_WITH_ACCESSIBLE_TRIP",
    "ALLOCATED_WAV"
  ),
  region     = region_name,
  services   = "TAXI",
  area_type  = "REGIONAL",
  date_from  = date_from,
  date_to    = date_to,
  title      = "Number of WAVs (vehicles)",

  colors = c(
    ACTIVE_WAV                      = "#69B7FF",
    ACTIVE_WAV_WITH_ACCESSIBLE_TRIP = "#1E3A8A",
    ALLOCATED_WAV                   = "#D55E00"
  ),

  # NEW: custom legend labels
  labels = c(
    ACTIVE_WAV                      = "Active WAVs",
    ACTIVE_WAV_WITH_ACCESSIBLE_TRIP = "Active WAVs Making Accessible Trips",
    ALLOCATED_WAV                   = "Board-Allocated WAVs"
  )
)

chart

```
Analysis and Conclusions

In `r end_label`, `r region_name` Regional District recorded a total of `r metric("Active WAVs","TAXI","last_value_fmt")` active WAVs. On a year-over-year (YoY) basis, the number of active WAVs in the taxi fleet was `r metric("Active WAVs","TAXI","yoy_fmt")`while on a year-to-date (YTD) basis it was `r metric("Active WAVs","TAXI","ytd_yoy_fmt")` relative to the same period in `r prev_year`. On a month-over-month (MoM) basis, the number of active WAVs in the taxi fleet changed by `r metric("Active WAVs","TAXI","mom_fmt")`. However, the number of active WAVs remained below the Board’s approved WAVs (`r metric("ALLOCATED_WAV","TAXI","last_value_fmt")` vehicles in `r end_label`), indicating that the full potential utilization of approved WAVs in this area had yet to be realized. `r trend_narrative("ALLOCATED_WAV","TAXI")` `r shift_narrative("ALLOCATED_WAV","TAXI")`

In `r end_label`, `r region_name` Regional District recorded a total of `r metric("ACTIVE_WAV_WITH_ACCESSIBLE_TRIP","TAXI","last_value_fmt")` active WAVs providing accessible trips. On a year-over-year (YoY) basis, the number of active WAVs providing accessible trips was `r metric("ACTIVE_WAV_WITH_ACCESSIBLE_TRIP","TAXI","yoy_fmt")`, while on a year-to-date (YTD) basis it was `r metric("ACTIVE_WAV_WITH_ACCESSIBLE_TRIP","TAXI","ytd_yoy_fmt")` relative to the same period in `r prev_year`. On a month-over-month (MoM) basis, the number of active WAVs providing accessible trips changed by `r metric("ACTIVE_WAV_WITH_ACCESSIBLE_TRIP","TAXI","mom_fmt")`. `r trend_narrative("ACTIVE_WAV_WITH_ACCESSIBLE_TRIP","TAXI")` `r shift_narrative("ACTIVE_WAV_WITH_ACCESSIBLE_TRIP","TAXI")`


<!-- In `r end_label`, `r region_name` Regional District recorded a total of `r metric("ACTIVE_WAV_PER_1000_PERSONS","TAXI","last_value_fmt")` vehicles per 1,000 persons. On a year-over-year (YoY) basis, the availability of active WAVs per 1,000 persons was `r metric("ACTIVE_WAV_PER_1000_PERSONS","TAXI","yoy_fmt")`, while on a year-to-date (YTD) basis it was `r metric("ACTIVE_WAV_PER_1000_PERSONS","TAXI","ytd_yoy_fmt")` relative to the same period in `r prev_year`. On a month-over-month (MoM) basis, the number of active WAVs per 1,000 persons changed by `r metric("ACTIVE_WAV_PER_1000_PERSONS","TAXI","mom_fmt")`. `r trend_narrative("ACTIVE_WAV_PER_1000_PERSONS","TAXI")` `r shift_narrative("ACTIVE_WAV_PER_1000_PERSONS","TAXI")` -->

<!-- In `r end_label`, `r region_name` Regional District recorded a total of `r metric("ACTIVE_WAV_WITH_ACCESS_TRIP_PER_1000_PERSONS","TAXI","last_value_fmt")` vehicles per 1,000 persons providing accessible trips. On a year-over-year (YoY) basis, the availability of active WAVs per 1,000 persons providing accessible trips was `r metric("ACTIVE_WAV_WITH_ACCESS_TRIP_PER_1000_PERSONS","TAXI","yoy_fmt")`, while on a year-to-date (YTD) basis it was `r metric("ACTIVE_WAV_WITH_ACCESS_TRIP_PER_1000_PERSONS","TAXI","ytd_yoy_fmt")` relative to the same period in `r prev_year`. On a month-over-month (MoM) basis, the number of active WAVs per 1,000 persons providing accessible trips changed by `r metric("ACTIVE_WAV_WITH_ACCESS_TRIP_PER_1000_PERSONS","TAXI","mom_fmt")`. `r trend_narrative("ACTIVE_WAV_WITH_ACCESS_TRIP_PER_1000_PERSONS","TAXI")` `r shift_narrative("ACTIVE_WAV_WITH_ACCESS_TRIP_PER_1000_PERSONS","TAXI")` -->


Limitations: The number of active WAVs does not capture the Board’s other aspects of accessibility, including accessibility for persons with other disabilities, overall availability of services, essential service levels, service to underserved groups or communities, and inclusion. 

