### Number of Active Vehicles per 1,000 Persons

Primary Factor: Accessibility

**Description:** The number of active vehicles per 1,000 persons represents the total number of vehicles serving every 1,000 persons in an area over a month.

**Measurement:** The number of active vehicles per 1,000 persons is calculated as the ratio of the total number of active vehicles (including both taxis and TNS) in an area relative to its population, normalized to a per 1,000 persons basis.

**Interpretation:** A higher number of active vehicles per 1,000 persons indicates higher accessibility to services in the area, as more taxis and TNS vehicles are available to serve every 1,000 persons.

Policy Context: Public Need and Accessibility Factor

The Board’s Public Need Policy in the **Policy Manual** lists examples of the accessibility factor, which are set out below alongside their corresponding implications on the number of active vehicles per 1,000 persons:

\-**The proposed service enhances availability of passenger transportation services to everyone in the province, including in low-density areas, such as rural and remote communities.** -A higher number of active vehicles per 1,000 persons directly and positively correlates with greater service availability. -The number of vehicles per 1,000 persons in low-density areas provides an area-specific assessment of overall service availability.

#### `r region_name` Regional District Active Vehicles per 1,000 Persons Findings

`r tbl_caption(paste(region_name, "Regional District Active Vehicles per 1,000 Persons"))`

```{r}
make_indicator_table_compact(
  data      = raw_norm,
  indicator = "VEHICLE_PER_1000_PERSONS",
  region    = region_name,
  date_from = date_from,
  date_to   = date_to,
  digits    = 2,
  currency  = FALSE)
```

Note: The above table reports the number of active vehicles per 1,000 persons up to the most recent year for which annual population data is available from Statistics Canada. As of October 2024, the most recent data available is up to 2023.[^4] 
[^4]: Statistics Canada. (2024). Table 17-10-0152-01 Population estimates, July 1, by census division, 2021 boundaries.https://www150.statcan.gc.ca/t1/tbl1/en/tv.action?pid=1710015201

`r fig_caption(paste(region_name, "Regional District Active Vehicles per 1,000 Persons"))`

```{r, fig.width=10, fig.height=5}
chart <- make_indicator_chart_stacked_2(
  data       = raw_norm,
  indicator  = "VEHICLE_PER_1000_PERSONS",
  region     = region_name,
  services   = c("TNS","TAXI"),   # first = bottom of stack
  area_type  = "REGIONAL",
  date_from  = date_from,
  date_to    = date_to,
  title      = "Active Vehicles per 1,000 Persons",
  caption    = NULL
)
chart

```

Analysis and Conclusions

In `r end_label`, `r region_name` Regional District had `r metric("VEHICLE_PER_1000_PERSONS","TAXI","last_value_fmt")` active taxi vehicles per 1000 persons. On a year-over-year (YoY) basis, the active taxi vehicles per 1000 persons changed by `r metric("VEHICLE_PER_1000_PERSONS","TAXI","yoy_fmt")`, while year-to-date (YTD), the figure was `r metric("VEHICLE_PER_1000_PERSONS","TAXI","ytd_yoy_fmt")` compared to the same period in `r prev_year`. On a month-over-month (MoM) basis, the active taxi vehicles per 1000 persons changed by `r metric("VEHICLE_PER_1000_PERSONS","TAXI","mom_fmt")`. `r trend_narrative("VEHICLE_PER_1000_PERSONS", "TAXI")` `r shift_narrative("VEHICLE_PER_1000_PERSONS", "TAXI")`

In `r end_label`, `r region_name` Regional District had `r metric("VEHICLE_PER_1000_PERSONS","TNS","last_value_fmt")` active TNS vehicles per 1000 persons. Compared with `r prev_label`, TNS active vehicles per 1000 persons changed by `r metric("VEHICLE_PER_1000_PERSONS","TNS","yoy_fmt")` (YoY). Year-to-date through `r end_label`, the active TNS vehicles per 1000 persons was `r metric("VEHICLE_PER_1000_PERSONS","TNS","ytd_yoy_fmt")` compared to the same period in `r prev_year`. On a month-over-month (MoM) basis, the active TNS vehicles per 1000 persons changed by `r metric("VEHICLE_PER_1000_PERSONS","TNS","mom_fmt")`. `r trend_narrative("VEHICLE_PER_1000_PERSONS", "TNS")` `r shift_narrative("VEHICLE_PER_1000_PERSONS", "TNS")`

**Limitations:** The number of active vehicles per 1,000 persons does not address the Board’s other examples of accessibility, including accessibility for people with other disabilities or underserved groups and communities. In addition, it does not reflect the reduction in the barriers to inclusion.
