### Number of Active Vehicles per 1,000 Persons

Primary Factor: Accessibility

**Description:**\
The number of active vehicles per 1,000 persons represents the total number of vehicles serving every 1,000 persons in an area over a month.

**Measurement:**\
The number of active vehicles per 1,000 persons is calculated as the ratio of the total number of active vehicles (including both taxis and TNS) in an area relative to its population, normalized to a per 1,000 persons basis.

**Interpretation:**\
A higher number of active vehicles per 1,000 persons indicates higher accessibility to services in the area, as more taxis and TNS vehicles are available to serve every 1,000 persons.

------------------------------------------------------------------------

## Policy Context: Public Need and Accessibility Factor

The Board’s *Public Need Policy* in the **Policy Manual** lists examples of the accessibility factor, which are set out below alongside their corresponding implications on the number of active vehicles per 1,000 persons:

-**The proposed service enhances availability of passenger transportation services to everyone in the province, including in low-density areas, such as rural and remote communities.**
    -A higher number of active vehicles per 1,000 persons directly and positively correlates with greater service availability.
    -The number of vehicles per 1,000 persons in low-density areas provides an area-specific assessment of overall service availability.

##### `r region_name` Regional District Active Vehicles per 1,000 Persons Findings
**Table 1: `r region_name` Regional District Active Vehicles per 1,000 Persons**

```{r}
source("setup.R")
```


```{r}
tbl <- make_indicator_table_compact(
  data       = raw_norm,
  indicator  = "VEHICLE_PER_1000_PERSONS",
  region     = region_name,
  services   = c("TAXI","TNS"),
  area_type  = "REGIONAL",
  date_from  = date_from,
  date_to    = date_to,
  digits     = 2
)
tbl
```

*Figure 1: `r region_name` Regional District Active Vehicles per 1,000 Persons*
```{r, fig.width=10, fig.height=5}
chart <- make_indicator_chart_stacked(
data       = raw_norm,
indicator  = "VEHICLE_PER_1000_PERSONS",
region     = region_name,
services   = c("TNS","TAXI"),  # first = bottom of stack
area_type  = "REGIONAL",
date_from  = date_from,
date_to    = date_to,
title      = " Active Vehicles per 1,000 Persons"
)
chart

```


```{r}


# metric("VEHICLE_PER_1000_PERSONS","TAXI","last_value_fmt")
metric <- function(ind, service = c("TAXI","TNS"), field = c("last_value_fmt","yoy_fmt","ytd_yoy_fmt")){
  service <- match.arg(service)
  field   <- match.arg(field)
  out <- tryCatch(metrics[[ind]][[service]][[field]], error = function(e) NA)
  if (is.null(out)) "n/a" else out
}

```
Analysis and Conclusions

In `r end_label`, `r region_name` Regional District saw a total of `r metric("VEHICLE_PER_1000_PERSONS","TAXI","last_value_fmt")` taxi trips revenue. On a year-over-year (YoY) basis, the total taxi trip revenue was `r metric("VEHICLE_PER_1000_PERSONS","TAXI","yoy_fmt")`, while on a year-to-date (YTD) basis it was `r metric("VEHICLE_PER_1000_PERSONS","TAXI","ytd_yoy_fmt")` relative to the same period in `r prev_year`. These results indicate strong momentum in demand for taxi services in the region.

In `r end_label`, `r region_name` Regional District recorded `r metric("VEHICLE_PER_1000_PERSONS","TNS","last_value_fmt")` TNS trips. Compared with `r prev_label`, TNS trip revenue was `r metric("VEHICLE_PER_1000_PERSONS","TNS","yoy_fmt")` (YoY). Year-to-date through `r end_label`, TNS trip revenue was `r metric("VEHICLE_PER_1000_PERSONS","TNS","ytd_yoy_fmt")` relative to the same period in `r prev_year`. These results indicate strong momentum in demand for TNS services in the region.


**Limitations:** The number of active vehicles per 1,000 persons does not address the Board’s other examples of accessibility, including accessibility for people with other disabilities or underserved groups and communities. In addition, it does not reflect the reduction in the barriers to inclusion.