## Innovation Factor (SEC)

There is no analysis related to this factor.

## Sustainability Factor (SEC)

### Revenue per Active Vehicle

Primary Factor: Sustainability

**Description:** Revenue per active vehicle measures the amount of revenue generated by each vehicle that was in operation in a given area during the reporting period.

**Measurement:** Revenue per active vehicle is calculated as the median (50th percentile) of the revenue-per-active-vehicle distribution in an area over a month.

**Interpretation:** Higher revenue per active vehicle supports the industry's sustainability by ensuring each vehicle generates more gross income.

Policy Context: Public Need and Accessibility Factor

The Board’s sound economic conditions policy in the Policy Manual lists examples of the sustainability factor, which are set out below alongside their corresponding implications on revenue per active vehicle:

-   The proposed service promotes long-term stability for the passenger transportation industry.
    -Assuming fixed costs, high and stable revenue per active vehicle implies high and stable profits, thus promoting long-term stability for the industry.
-   The proposed service encourages resiliency in the passenger transportation industry.
    -Assuming fixed costs, high and stable revenue per active vehicle implies high and stable profits, thus encouraging industry resilience.
-   The proposed service minimizes significant market disruption.
    -Assuming fixed costs, stable or gradually increasing revenue per active vehicle in the taxi sector indicates a lack of rapid disruption.
-   The proposed service promotes overall profitability and economic sustainability of the industry.
    -Assuming fixed costs, stable or gradually increasing revenue per active vehicle in the taxi sector indicates overall profitability and economic sustainability for the industry.

#### `r region_name` Regional District Revenue per Active Vehicle Findings


`r tbl_caption(paste(region_name, "Regional District Revenue per Active Vehicle"))`


```{r}
tbl <- make_indicator_table_compact(
  data       = raw_norm,
  indicator  = "REVENUE_PER_VEHICLE",
  region     = region_name,
  services   = c("TAXI","TNS"),
  area_type  = "REGIONAL",
  date_from  = date_from,
  date_to    = date_to,
  digits     = 2
)
tbl
```


`r fig_caption(paste(region_name, "Regional District Revenue per Active Vehicle"))`

```{r, fig.width=10, fig.height=5}
chart <- make_indicator_chart_lines(
data       = raw_norm,
indicator  = "REVENUE_PER_VEHICLE",  # Adjust indicator if needed
region     = region_name,
services   = c("TAXI", "TNS"),
area_type  = "REGIONAL",
date_from  = date_from,
date_to    = date_to,
title      = "Revenue per Active Vehicle",
colors     = c(TAXI = "#69B7FF", TNS = "#1E3A8A")
)
chart

```



Analysis and Conclusions

In `r end_label`, `r region_name` Regional District recorded a total of `r metric("REVENUE_PER_VEHICLE","TAXI","last_value_fmt")` in revenue per active vehicle for taxi trips. On a year-over-year (YoY) basis, the revenue per active vehicle for taxi trips was `r metric("REVENUE_PER_VEHICLE","TAXI","yoy_fmt")`while on a year-to-date (YTD) basis it was `r metric("REVENUE_PER_VEHICLE","TAXI","ytd_yoy_fmt")` relative to the same period in `r prev_year`. On a month-over-month (MoM) basis, revenue per active taxi vehicle changed by `r metric("REVENUE_PER_VEHICLE","TAXI","mom_fmt")`. `r trend_narrative("REVENUE_PER_VEHICLE", "TAXI")` `r shift_narrative("REVENUE_PER_VEHICLE", "TAXI")`

In `r end_label`, the `r region_name` Regional District recorded a total of `r metric("REVENUE_PER_VEHICLE","TNS","last_value_fmt")` in revenue per active vehicle for TNS trips. Compared with `r prev_label`, the revenue per active vehicle for TNS trips was `r metric("REVENUE_PER_VEHICLE","TNS","yoy_fmt")` (YoY). Year-to-date through r end_label, the revenue per active vehicle for TNS trips was `r metric("REVENUE_PER_VEHICLE","TNS","ytd_yoy_fmt")` relative to the same period in `r prev_year`. On a month-over-month (MoM) basis, revenue per active TNS vehicle changed by `r metric("REVENUE_PER_VEHICLE","TNS","mom_fmt")`. `r trend_narrative("REVENUE_PER_VEHICLE", "TNS")` `r shift_narrative("REVENUE_PER_VEHICLE", "TNS")`

**Limitations:** The median (i.e., the 50th percentile) revenue per active vehicle cannot fully represent the sustainability of all active vehicles in a region in a month. Moreover, revenue per active vehicle does not address the Board’s other sustainability examples, including balance across different sectors, promotion of driver incomes and contribution to a stable supply of drivers in the taxi sector, environmental sustainability, reduction of greenhouse gas emissions, other pollutions, congestion, or reduction of harm to or enhancement of Public Transportation.
