### Taxi Fleet Utilization Rate

**Primary factor:** Sustainability

**Description:** Taxi fleet utilization rate represents the proportion of licensed vehicles that are active in an area over a month.

**Measurement:** The taxi fleet utilization rate is determined by dividing the number of active vehicles by the total number of licensed vehicles in a given area during the reporting period. The number of active vehicles is determined by data reported to the Trip Database, while the maximum fleet size is based on the 

**Interpretation:** A relatively stable taxi fleet utilization rate over time can indicate long-term stability for the taxi and TNS industry . On the other hand, a decline in utilization rates, other things being equal, suggests unsustainable economic conditions as licensees choose to operate fewer vehicles in their fleet.

The Board’s sound economic conditions policy in the Policy Manual lists examples of the sustainability factor, which are set out below alongside their corresponding implications on taxi fleet utilization rate:

-   The proposed service promotes long-term stability for the passenger transportation industry.

    -   Stability in taxi fleet utilization rate can indicate the long-term stability of the industry.

-   A higher taxi fleet utilization rate indicates greater resiliency in the passenger transportation industry.

    -   A consistently close-to-full fleet utilization rate indicates the industry is at its supply capacity, and unable to meet further demand growth. A consistently low fleet utilization can indicate a lack of growth of the industry.

-   The proposed service minimizes significant market disruption.

    -   A low taxi fleet utilization rate increases the risk of market oversupply.

-   The proposed service protects the taxi sector from rapid disruption to ensure its long-term stability as an essential service.

    -   A low taxi fleet utilization rate increases the risk of rapid disruption in the taxi sector.

-   The proposed service promotes overall profitability and economic sustainability of the industry.

    -   A low taxi fleet utilization rate suggests lower profitability or economic unsustainability in the market. 

#### `r region_name` Regional District Taxi Fleet Utilization Rate Findings


`r tbl_caption(paste(region_name, "Regional District taxi fleet utilization rate"))`


```{r}
make_indicator_table_compact(
  data      = raw_norm,
  indicator = "FLEET_UTILIZATION_RATE",
  region    = region_name,
  date_from = date_from,
  date_to   = date_to,
  digits    = 2,
  currency  = FALSE)
```


`r fig_caption(paste(region_name, "Regional District taxi fleet utilization rate"))`

```{r, fig.width=10, fig.height=5} 
chart <- make_indicator_chart_lines(
data       = raw_norm,
indicator  = "FLEET_UTILIZATION_RATE",  # Adjust indicator if needed
region     = region_name,
services   = c("TAXI", "TNS"),
area_type  = "REGIONAL",
date_from  = date_from,
date_to    = date_to,
title      = "Taxi Fleet Utilization Rate",
colors     = c(TAXI = "#69B7FF", TNS = "#1E3A8A")
)
chart

```



Analysis and Conclusions

In `r end_label`, `r region_name` Regional District recorded a total of `r metric("FLEET_UTILIZATION_RATE","TAXI","last_value_fmt")` in taxi fleet utilization rate  trips. On a year-over-year (YoY) basis, the taxi fleet utilization rate  trips was `r metric("FLEET_UTILIZATION_RATE","TAXI","yoy_fmt")`while on a year-to-date (YTD) basis it was `r metric("FLEET_UTILIZATION_RATE","TAXI","ytd_yoy_fmt")` relative to the same period in `r prev_year`. On a month-over-month (MoM) basis, taxi fleet utilization changed by `r metric("FLEET_UTILIZATION_RATE","TAXI","mom_fmt")`. `r trend_narrative("FLEET_UTILIZATION", "TAXI")` `r shift_narrative("FLEET_UTILIZATION", "TAXI")`


**Limitations:** The taxi fleet utilization rate does not provide insight into the equitability of opportunities in the industry, the impact on Public Transportation, the industry's sustainable profitability levels, or driver incomes, nor does it address the environmental aspects of sustainability.
